### 2024-07-04 thesis/ms-fletcher-porter.tex
Edit pass for own comments

\usepackage{siunitx}
\usepackage{gensymb}
\usepackage{svg}
\usepackage{lipsum}  % for placeholder text
\usepackage{subcaption}
\usepackage{bigfoot} % for \verb in \footnote

@@

An important property of layered abstraction architectures is that protocols which are defined in higher layers can have the lower layers transparently replaced. For instance, the internet protocol (IP) which we use every day is defined in the Network layer. Typically, the internet is built over Ethernet or Wi-Fi, but with some effort, other low-level protocols are available. Lindgren et al.\ developed IP over CAN, wherein an internet network is extended over a CAN network~\cite{Lindgren2008}, which was never imagined by implementers of CAN nor IP.

@@

The OSI standard also includes the principles upon which its layers were designed~\cite[§6.2.1]{ISO7498-1}. In summary, each layer should be a discrete, separable item with a coherent function. Their interfaces should be succinct and only connect with adjacent layers. Fewer layers eases explanation and implementation and the choice layers should be motivated by experience. Being a novel theory, experience will be the principal gain of future work, although case studies discussed later will also help. Usefully, it offers a test to aid in determining layer boundaries. Layers should have a different level of abstraction in terms of \textit{morphology}, \textit{syntax}, and \textit{semantics}.

@@

Morphology is concerned with the atomic elements of meaning. In C code, this means short tokens of text, here \verb|printf|, \verb|(|, \verb|"I have %u cows\n"|, \verb|,|, \verb|num_cows|, \verb|)|, and \verb|;|. None of these can be reduced and still carry meaning---\verb|pri| and \verb|ntf| doesn't make \verb|printf|. But these atoms of meaning needn't be text or language. As will be discussed, they can be cables, connectors, computer hardware, or anything with relational meaning.

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier \verb|printf| has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of these tokens mean is the domain of semantics.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/printf_syntax}
  \caption{The above printf example with syntax labeled according to K{\&}R C~\cite{Kernighan1978}}\label{fig:printf_syntax}
\end{figure}

To understand the semantics of \verb|printf|, you must read the documentation. It will tell you that \verb|printf| means to convert and write output to the console under the control of a format string. You'll also see that \verb|"I have %u cows\n"| is that format string, and \verb|num_cows| is an additional argument, presumably a number of cows. Reading the documentation further, you'll see that the format string itself has its own morphology, syntax, and semantics driving how output is created.

@@

Next, these ideas will be used to make a mechatronic tech stack like the web and OSI stacks.

@@

Morphology provides the strongest and most coarse distinctions in layers. If the atoms that one component is made of are incompatible with another, they can't effectively be considered together. The most obvious case is that those components that are made of literal atoms, i.e.\ hardware, must be considered separately from those that aren't physical, namely software. Within hardware, a control engineer will concern themselves with the components of the computer which their control software run on (RAM, CPU, storage, etc.) and the physical connections between the computer components and the sensors and transducers, such as cables, connectors, contacts, and traces. The software is the instructions of the control program itself, but also supporting software like the operating system, runtime environments, and libraries.

Syntax is the structure of an idea. It is a finer comb to discriminate layers. Software generally is built as a linear sequence of instructions with branches and loops. This describes the control engineer's perspective on their own control program, but supporting software like libraries are more abstract. They're better thought of as a dependency. A control program may depend on libraries A, B, and C, and those in turn depend on libraries D, E, F, and G. These \textit{middlewares} are better understood as a network of the programs themselves and the dependency connections between them. For hardware, its structure is its assemblage. The computer components are connected to the motherboard, sensors and transducers are connected to a CAN bus, the temperature sensors message over I2C, etc. Like middleware, hardware is the network of components and the connections between them, but the type of connection is also important. A system with components on a CAN bus is going to be different than one with a USB bus, and different still from a system where all components communicate over a serial connection.

@@

This is enough to propose a mechatronic tech stack. The mechatronic stack as outlined sits on top of signaling with sensors and transducers. Those are interfaced by IO systems that connect to computer systems which run middleware upon which a control program depends. This is illustrated as a stack diagram in figure~\ref{fig:mechatronic_tech_stack}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/mechatronic_tech_stack}
  \caption{The mechatronic tech stack places high-level business logic at the top and low-level interfacing at the bottom.}\label{fig:mechatronic_tech_stack}
\end{figure}

@@

These case studies will apply the theory of a mechatronic tech stack to three projects---one where a requirement change caused a re-architecture of the control system, another where three similar devices have their stacks compared, and the last has its stack designed with this theory in mind. In all cases, the project's stack will be explicated and there will be a discussion of interesting features of its stack.

@@

Dolores is a full-scale test bed in Aalto University's Fluid Power Laboratory, seen in figure~\ref{fig:dolores0} to study the application of electric hydrostatic actuation in an articulated-steering linkage for a front-loading mining vehicle~\cite{Hermansson2021, Helduser1999, Dudzinski1989}. Its control system commissioning was completed in 2021, but come 2023, an industry partner in the experiment requested hardware changes that demanded a substantial re-architecture of the control system. This case study will discuss the mechatronic tech stacks behind the old and the new systems.

@@

The industry partner's requested change occurred in the IO (input/output) layer. The old system had an inverter that drives a servomotor, a safety valve to enable the system, and an array of pressure, temperature, and flow sensors to observe the state of the hydraulic system. The inverter and the pressure and temperature sensors all communicated over the CANopen protocol, the flow sensors gave analog inputs, and the safety valve was actuated with a digital signal. The CAN communication was interfaced by a Vector CAN/USB dongle~\cite{VectorCanUsbDongle}. All other IO went through a National Instruments data acquisition device (DAQ). In the new system, the forced change for the inverter to communicate over EtherCAT.\ This meant new IO hardware was needed to support this protocol, which was selected to be Beckhoff EtherCAT terminals. They provided the interface to connect the inverter, digital and analog IO, and a CAN bus for the sensor network~\cite{BeckhoffEL6751}. This made the DAQ unnecessary, but the CAN/USB dongle was still required due to deficiencies in Beckhoff's control interfaces, which are discussed with the control layer.

The system had been controlled by a conventional Windows workstation. This may have been fine to keep for the new system, but there was concern with using a whole workstation for a single experiment. Instead, one of Beckhoff's industrial PCs was selected as the computer to leverage Beckhoff's support~\cite{BeckhoffC60XX}. It was expensive, but it was thought that architecting the system to align with vendor expectations would make development easier and mitigate potential problems. However, the workstation did tap into the CAN bus to initialize the sensors, so it wasn't completely out of the loop.

In the old system, the workstation ran Windows, the DAQ was managed by drivers from National Instruments, and the CAN bus was initialized using Vector's CANalyzer application and runtime commands came through Vector's Simulink library~\cite{VectorCanalyzer, SimulinkVectorCAN}. Due to issues getting reliable EtherCAT connections with Simulink (possibly for user error, though the documentation wasn't simple), the new system uses a PLC-based middleware. Still based on Windows, it has libraries and runtimes to run the EtherCAT bus, to read from the CAN bus, and to interface with an external development computer. CANalyzer remained to initialize the CAN bus.

The old system had controllers written with Matlab and Simulink. However, due to middleware issues, they were abandoned in favor of writing controllers in PLC languages based on IEC~61131-3 with the TwinCAT development environment~\cite{IEC61131-3, BeckhoffTwinCAT}. The major issue with TwinCAT is that its CANopen implementation had no documented way to initialize devices. This was worked around by initializing devices with CANalyzer on the workstation.

@@

Notably, a small change in the system, a change in IO protocol, demanded a wholesale re-architecture of the control system. Also notable is that issues in controller-layer tooling echoed through the whole stack. This speaks to how in both stacks, the tools are \textit{tightly coupled}, meaning changes in one tool often demand changes in the others. For this mechatronic stack model to be a useful means of selecting tools like the web stack, the layers need to be \textit{loosely} coupled so tools can be replaced with minimal change in other layers.

@@

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an EFPA~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work I did not participate in, some details must be inferred.

\begin{figure}[h]
  \centering
  \begin{subfigure}[t]{0.47\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/six_legged_robot}
    \caption{Six-legged robot~\cite{Hase2020}}\label{sfig:six_legged_robot}
  \end{subfigure}
  \begin{subfigure}[t]{0.51\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/pipe_robot}
    \caption{Pipe-traversal robot~\cite{Shinohara2020}}\label{sfig:pipe_robot}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/bending_servo_valve}
    \caption{Bending servo valve~\cite{Kobayashi2020}}\label{sfig:bending_servo_valve}
  \end{subfigure}
  \caption{These projects are made by the same people with the same technologies using similar tools.}\label{fig:soft_pneumatic_actuators_demo}
\end{figure}

In the IO layer, all three are similar. The six-legged and pipe robots have an array of on/off valves commanded by digital signals to pressurize their EFPAs. The servo valve has its position set with a small motor, which is controlled by a PWM signal. These signals are propagated through direct connection to pins on the computer, although Shinohara et al.\ note that there are circuits between stepping up power. The robots received command signals from a workstation over a serial connection. Hase et al.\ note that there was an intermediate chip on the six-legged robot to connect over USB.\ The servo valve computer had no reported inputs, though several sensors were connected to a data recording device.

The computer used for the two robots was an SH7125, a 32-bit SuperH-based system on a chip, a device that has all necessary components for computation like CPU, RAM, and peripherals on a single die~\cite{SH7125}. The servo valve used an H8/3264, a 32-bit H8/300-based system on a chip~\cite{H83264}. These architectures, while unfamiliar in name, are common for applications like printer or hard drive controllers where cost and hardware IO are important. Both CPUs have inbuilt peripherals for IO where when the CPU writes a number to a register, the IO pins will change state to match its binary representation.

No middleware is mentioned in these papers. While it's in principle possible that they ran some embedded operating system, it's doubtful given all their controllers seem fairly simple, as discussed below. It is possible that they used some external libraries to aid in programming. A programming environment without an operating system is referred to as \textit{bare metal} in industry parlance because that's all there is to the middleware interface.

The language the controllers were written in was not mentioned in the papers, but it's likely they used an assembly language or C since major compilers support both of these architectures~\cite{GccSupportedArchitectures}. Hase et al.\ state that the control program they wrote was to cycle through a fixed sequence of states for the valves, perhaps switching sequences interactively over the serial connection.

Each of these three projects implemented a largely similar stack---IO based on direct connections to the CPU with supporting circuitry, an embedded system-on-a-chip computer, no middleware, and controllers written in C or assembly. By using a common stack, these researchers were able to skip over the tedious and difficult work of selecting tools and just use what is familiar to their institution and focus their energy on research.

@@

With this new theory of a mechatronic tech stack, it made sense to start a project to study what tools this model lends itself to when applied prescriptively. The project was a driving simulator chair that had been an unfinished student project a year prior~\cite{Bjoerklund2023}. The mechanical systems had been finished including all the sensors and transducers, so all that was needed was a control system.

@@

The system is a chair atop a Stewart platform whose linear actuators are single-acting pneumatic-muscle cylinders, seen in figure~\ref{fig:stewart_platform_demo}~\cite{Stewart1965, Caldwell1995}. The valve manifolds that were already made for this had one valve to connect the pneumatic muscle to an external pressure source and another to vent the muscle to atmosphere. When both valves are closed, it brakes the cylinder. There are six instances of this circuit, connected in series to a common pressure source. Each cylinder also has a position sensor exposed as a voltage signal from a potentiometer.

Tool selection started from the IO layer. The need was to switch 12 digital outputs for the valves and six analog inputs ($\qtyrange[range-units=single,range-phrase=..]{0}{24}{\volt}$) from the position sensors. It was also desirable for the work to be done quickly to have a demonstration in time for publication. This was met with EtherCAT IO cards from Beckhoff~GmbH, in large part thanks to their availability within our lab. They supply IO cards that mostly met the requirements; the digital outputs needed to be supplemented with flyback diodes for the inductive load and the analog inputs needed signals stepped down to $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ with resistors~\cite{BeckhoffEL2042,BeckhoffEL3062}. This meant routing signal cables from the sensors and transducers to terminal blocks, placing circuitry in the terminals, and then routing to the Beckhoff terminals, where there is an EtherCAT connection to the computer's Ethernet port~\cite{BeckhoffEK1100}.

In selecting the computer, it needed to communicate with the IO over EtherCAT and be cheap and easy to develop for. The Raspberry Pi supports EtherCAT, is reasonably priced, and Linux-based systems are familiar to the authors and easy to develop software for.

The principal middleware requirement was an EtherCAT master, which itself depends on system-level networking. Linux satisfies the networking requirement. EtherCAT masters are conventionally provided by PLC-based systems, like those from Beckhoff~\cite{BeckhoffTwinCAT}. However, they generally do not run on Linux. However, the point of this project was to try new things. So, Ethercrab, a master written in the Rust programming language, was used~\cite{Ethercrab}. Ethercrab benefits from being open source, so motivated users can troubleshoot more easily. It also has thorough documentation and lots of examples (e.g.\ the control-layer application was based on \verb|ethercrab/examples/ek1100.rs|). Interfacing with Rust is a benefit due to its ease of use, quality tooling, performance, and memory safety~\cite{RustHomePage}.

Finally, the control layer had to use the API of the middleware layer to implement the business logic. One case of that logic was to have a controller to regulate the position of each cylinder. It had to be implemented in Rust to interface with the Ethercrab library. It also required some code to initialize the EtherCAT communication and to configure the Beckhoff terminals. There's also a user interface to interactively set the cylinder position. An interesting issue came up where the position sensor signal, mapping the $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ IO signal to the range 0..1~\verb|float|, but only reached the range 0..0.5~\verb|float|. This was because the step-down resistor was larger than designed since the right size wasn't available, so the maximum sensor reading was smaller than $\qty{10}{\volt}$. This unhandled gain is an undesired interface between the Control and IO layers since it demands the controller be written with knowledge of the minutia of how the IO was implemented. Further research will go toward how to prevent this.

Also interesting in the architecture of a control system is what development environments can be constructed for it.  The development environment is distinct from the heretofore discussed mechatronic stack since it is not a part of the function of the running system, but instead is means which an engineer uses to turn an idea into assets for a mechatronic system.  For a given tech stack, any development environment generally can in principle yield the same assets, but a good tech stack may lend itself to higher quality, more consistent output, faster development cycles, or at least to be enjoyable for individual engineers to use.

>>>>

\usepackage{siunitx}
\usepackage{gensymb}
\usepackage{svg}
\usepackage{lipsum}  % for placeholder text
\usepackage{subcaption}
\usepackage{bigfoot} % for \verb in \footnote
\usepackage{cprotect} % for \verb in \caption

@@

An important property of layered abstraction architectures is that protocols which are defined in higher layers can have the lower layers transparently replaced. For instance, the internet protocol (IP) which we use every day is defined in the Network layer. Typically, the internet is implemented on top of Ethernet or Wi-Fi, but with some implementation, other low-level protocols are available. Lindgren et al.\ developed IP over CAN, wherein an internet network is extended over a CAN network~\cite{Lindgren2008}, which was never imagined by the implementers of CAN nor IP.

@@

The OSI standard also includes the principles upon which its layers were designed~\cite[§6.2.1]{ISO7498-1}. In summary, each layer should be a discrete, separable item with a coherent function. Their interfaces should be succinct and only connect with adjacent layers. Fewer layers eases explanation and implementation and the choice of layers should be motivated by experience. This being a novel theory, experience will be the principal gain of future work, although case studies discussed later will also help. Helpfully, it offers a test to aid in determining layer boundaries. Layers should have a different level of abstraction in terms of \textit{morphology}, \textit{syntax}, and \textit{semantics}.

@@

Morphology is concerned with the atomic elements of meaning. In C code, this means short tokens of text, here \verb|printf|, \verb|(|, \verb|"I have %u cows\n"|, \verb|,|, \verb|num_cows|, \verb|)|, and \verb|;|. None of these can be reduced and still carry meaning---\verb|pri| and \verb|ntf| seperately don't combine to make \verb|printf|. But these atoms of meaning needn't be text or language. As will be discussed, they can be cables, connectors, computer hardware, or anything that carries semantic meaning.

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier \verb|printf| has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of this means is the domain of semantics.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/printf_syntax}
  \cprotect\caption{The above printf example with syntax labeled according to Kernighan {\&} Ritchie C~\cite{Kernighan1978}.  The whole line is a postfix expression and the \verb|printf| token is also a postfix expression on its own.  The string between parentheses is an argument expression list, and the strings between commas are a variety of syntactic expressions that enable writing non-trivial code, terminating as a primary expression, and then here has a string literal and an identifier.}\label{fig:printf_syntax}
\end{figure}

The semantics of \verb|printf| are the exprected behavior of the function, which can be gleaned from the documentation. It will say that \verb|printf| means to convert and write output to the console under the control of a format string~\cite{Kernighan1978}.  It will also say that the format string is the first expression in th argument expression list and that it is of type \verb|char*|.  So \verb|"I have %u cows\n"| is that format string, and the documentation will state that the second expression in the argument expression list, here \verb|num_cows| will be formatted as a string and written to the output where conversion character, here \verb|%u| is.  The logic of how the conversion characters and string formatting work are based on the format string's own morphology, syntax, and semantics, the study of which is left as an exercise to the reader.

@@

Next, these ideas will be used to make a mechatronic tech stack which combines the tool selecting power of the web stack and the abstract behavior of the OSI stack.

@@

Morphology provides the strongest and most coarse distinctions in layers. If the atomic components of one object aren't comparable to another, they can't usefully be considered together, like comparing apples and oranges. The most obvious case is that those components that are made of literal atoms, i.e.\ hardware, must be considered separately from those that aren't physical, namely software. Within hardware, a control engineer will concern themselves with the components of the computer which their control software run on (RAM, CPU, storage, etc.) and the physical connections between the computer components and the sensors and transducers, such as cables, connectors, contacts, and traces. The software is the instructions of the control program itself, but also supporting software like the operating system, runtime environments, and libraries.

Syntax is the structure of an idea. It is a finer comb to discriminate layers. Software generally is built as a linear sequence of instructions with branches and loops. This describes the control engineer's perspective on their own control program, but supporting software like libraries are more abstract. They're better thought of as a dependency. A control program may depend on libraries A, B, and C, and those in turn depend on libraries D, E, F, and G. These \textit{middlewares}, named so as they exist in the middle between hte control software and the computer, are better understood as a network of the programs themselves and the dependency connections between them. For hardware, its structure is its assemblage. The computer components are connected to the motherboard, sensors and transducers are connected to a CAN bus, the temperature sensors message over I2C, etc. Like middleware, hardware is a network of components and the connections between them, but the type of connection is also important. A system with components on a CAN bus is going to be different than one with a USB bus, and different still from a system where all components communicate over a serial connection.

@@

This is enough to propose a mechatronic tech stack. The mechatronic stack as outlined sits on top of signaling with sensors and transducers. Those are interfaced by IO systems that connect to computer systems which run middleware upon which a control program depends. The design is so that functionality in the control layer should be preserved if components in the lower layers are changed, perhaps wih some implementation effort, as was the case for IP over CAN.  This is illustrated as a stack diagram in figure~\ref{fig:mechatronic_tech_stack}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/mechatronic_tech_stack}
  \caption{The mechatronic tech stack places high-level business logic at the top and low-level interfacing at the bottom.  By siloing these function into seperate layers, their behaviors and constraints can be designed seperately.}\label{fig:mechatronic_tech_stack}
\end{figure}

@@

These case studies will apply the theory of a mechatronic tech stack to four projects---one where a requirement change caused a re-architecture of the control system, another where three similar devices have their stacks compared, the third has its stack designed with this theory in mind, and the last considers the design of a control system against challenging requirements. In all cases, the project's stack will be explicated and there will be a discussion of interesting features of its stack.

@@

Dolores is a full-scale test bed in the Aalto University Fluid Power Laboratory, seen in figure~\ref{fig:dolores0} to study the application of electric hydrostatic actuation in an articulated-steering linkage for a front-loading mining vehicle~\cite{Hermansson2021, Helduser1999, Dudzinski1989}. Its control system commissioning was completed in 2021, but in 2023, an industry partner in the experiment requested hardware changes that demanded a substantial re-architecture of the control system. This case study will discuss the mechatronic tech stacks behind the old and the new systems.

@@

The industry partner's requested change occurred in the IO (input/output) layer. The system has an inverter that drives a servomotor, a safety valve to enable the system, and an array of pressure, temperature, and flow sensors to observe the state of the hydraulic system. In the originally commissioned system, the inverter and the pressure and temperature sensors all communicated over a CANopen bus, the flow sensors gave analog inputs, and the safety valve was actuated with a digital signal. The CAN communication was executed by a Vector CAN/USB dongle~\cite{VectorCanUsbDongle}. All other IO went through a National Instruments data acquisition device (DAQ). In the new system, the forced change was for the inverter to communicate over EtherCAT instead of over CANopen.  The existing instrumentation didn't support EtherCAT, so this meant new IO hardware was needed to support this protocol, which was selected to be Beckhoff EtherCAT terminals. They provided the interface to connect the inverter, digital and analog IO, and a CAN bus for the sensor network~\cite{BeckhoffEL6751}. This made the DAQ unnecessary, but the CAN/USB dongle was still required due to deficiencies in Beckhoff's control interfaces that were discovered during development, which are discussed with the control layer.

The system had been controlled by a conventional Windows workstation. This may have been fine to keep for the new system, but there was concern with using a whole workstation only for a single experiment. Instead, one of Beckhoff's industrial PCs was selected as the computer to match the IO terminals and so leverage Beckhoff's support~\cite{BeckhoffC60XX}. It was expensive, but it was thought that architecting the system to align with vendor expectations would make development easier and mitigate potential problems. However, the desktop workstation did tap into the CAN bus to initialize the sensors, so it wasn't completely out of the loop.

In the old system, the workstation ran Windows, the DAQ was managed by drivers from National Instruments, and the CAN bus was initialized using Vector's CANalyzer application and runtime commands came through Vector's Simulink library~\cite{VectorCanalyzer, SimulinkVectorCAN}. Due to issues getting reliable EtherCAT connections with Simulink (possibly for user error, though the documentation wasn't clear), the new system uses a PLC-based middleware. Still based on Windows, it has libraries and runtimes to run the EtherCAT bus, to read from the CAN bus, and to interface with an external development workstation. CANalyzer remained to initialize the CAN bus.

The old system had controllers written with Matlab and Simulink. However, due to middleware issues, they were abandoned in favor of writing controllers in PLC languages based on IEC~61131-3 with the TwinCAT development environment~\cite{IEC61131-3, BeckhoffTwinCAT}. The major issue with TwinCAT is that its CANopen implementation has no documented way to initialize devices on startup. This was worked around by initializing devices with CANalyzer on the workstation.

@@

Notably, a small change in the system, a change in IO protocol, demanded a wholesale re-architecture of the control system. Also notable is that the CAN issues in controller-layer tooling echoed through the whole stack. This speaks to how in both stacks, the tools are \textit{tightly coupled}, meaning changes in one tool often demand changes in the others. For this mechatronic stack model to be a useful means of selecting tools like the web stack, the layers need to be \textit{loosely} coupled so tools can be replaced with minimal change in other layers.

@@

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an extension-type flexible pneumatic actuator (EFPA)~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work I did not participate in, some details are not documented and must be inferred.

\begin{figure}[h]
  \centering
  \begin{subfigure}[t]{0.47\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/six_legged_robot}
    \caption{Six-legged robot~\cite{Hase2020}}\label{sfig:six_legged_robot}
  \end{subfigure}
  \begin{subfigure}[t]{0.51\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/pipe_robot}
    \caption{Pipe-traversal robot~\cite{Shinohara2020}}\label{sfig:pipe_robot}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/bending_servo_valve}
    \caption{Bending servo valve~\cite{Kobayashi2020}}\label{sfig:bending_servo_valve}
  \end{subfigure}
  \caption{These projects are made by the same people with the same technologies using similar tools.  Photos taken from the cited papers.}\label{fig:soft_pneumatic_actuators_demo}
\end{figure}

In the IO layer, all three are similar. The six-legged and pipe robots have an array of on/off valves commanded by digital signals to pressurize their EFPAs. The servo valve has its position set with a small motor, which is controlled by a PWM signal. These signals are propagated through direct connection to pins on the computer, although Shinohara et al.\ note that there are circuits between stepping up power~\cite{Shinohara2020}. The robots received command signals from a workstation over a serial connection. Hase et al.\ note that there was an intermediate chip on the six-legged robot to connect over USB~\cite{Hase2020}. The servo valve computer had no reported inputs, though several sensors were connected to a data recording device.

The computer used for the two robots was an SH7125, a 32-bit SuperH-based system on a chip, a device that has all necessary components for computation like CPU, RAM, and peripherals on a single die~\cite{SH7125}. The servo valve used an H8/3264, a 32-bit H8/300-based system on a chip~\cite{H83264}. These architectures, while unfamiliar in name, are common for applications like printer or hard drive controllers where cost and hardware IO are important. Both CPUs have inbuilt peripherals for IO where when the CPU writes a number to a register, the IO pins will change state to match the IO register's binary representation.

No middleware is mentioned in these papers. While it's in principle possible that they ran some embedded operating system, it's doubtful given all their controllers are quite simple, as discussed below. It is possible that they used some external libraries to aid in programming.  In industry parlance, this kind of system without an operating system is a \textit{bare metal} system; so-called because that's all there is to the middleware interface.

The language the controllers were written in was not mentioned in the papers, but it's likely they used an assembly language or C since major compilers support both of these architectures~\cite{GccSupportedArchitectures}. Hase et al.\ state that the control program they wrote was to cycle through a fixed sequence of states for the valves, perhaps switching sequences interactively over the serial connection~\cite{Hase2020}.  This would be a controller based on a finite-state machine.

Each of these three projects implemented a largely similar stack---IO based on direct connections to the CPU with supporting circuitry, an embedded system-on-a-chip computer, bare metal middleware, and controllers written in C or assembly. By using a common stack, these researchers were able to skip over the tedious and difficult work of selecting tools and just use what is familiar to their institution and focus their energy on research.

@@

In claiming that this mechatronic tech stack aids development, it would be prudent to start a project to study what tools this model lends itself to when applied prescriptively. The project was a driving simulator chair that had been an unfinished student project a year prior~\cite{Bjoerklund2023}. The mechanical systems had been finished including all the sensors and transducers, so all that was needed was a control system.

@@

The system is a chair atop a Stewart platform whose linear actuators are single-acting pneumatic-muscle cylinders, seen in figure~\ref{fig:stewart_platform_demo}~\cite{Stewart1965, Caldwell1995}. The valve manifolds that were already made for this had one valve to connect the pneumatic muscle to an external pressure source and another to vent the muscle to atmosphere. When both valves are closed, the cylinder brakes. There are six instances of this circuit, connected in series to a common pressure source. Each cylinder also has a position sensor exposed as a voltage signal from a potentiometer.

Tool selection started from the IO layer. The need was to switch 12 digital $\qty{24}{\volt}$ outputs for the valves and six analog inputs ($\qtyrange[range-units=single,range-phrase=..]{0}{24}{\volt}$) from the position sensors. It was also desirable for the work to be done quickly to have a demonstration in time for publication. This was met with EtherCAT IO cards from Beckhoff~GmbH, in large part thanks to their availability within the Aalto University Fluid Power Laboratory. They supply IO cards that mostly met the requirements; the digital outputs needed to be supplemented with flyback diodes to handle the inductive load and the analog input signals needed to be stepped down to $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ with resistors~\cite{BeckhoffEL2042,BeckhoffEL3062}. This meant routing signal cables from the sensors and transducers to terminal blocks, placing circuitry in the terminals, and then routing to the Beckhoff terminals, where there is an EtherCAT connection to the computer's Ethernet port~\cite{BeckhoffEK1100}.

In selecting the computer, it needed to communicate with the IO over EtherCAT and be cheap and easy to develop for. The Raspberry Pi supports EtherCAT, is reasonably priced, and Linux-based systems are familiar and easy to develop software for.  The simplicity of the controller that was implemented also didn't demand much from the computer.

The principal middleware requirement was an EtherCAT master, which itself depends on system-level networking. Linux satisfies the system-level networking requirement. EtherCAT masters are conventionally provided by PLC-based systems, like those from Beckhoff~\cite{BeckhoffTwinCAT}. However, they generally do not run on Linux. Additionally, their behavior is opque and does not lend itself to understanding the system. So, Ethercrab, a master written in the Rust programming language, was used~\cite{Ethercrab}. Ethercrab benefits from being open source, so motivated users can troubleshoot more easily. It also has thorough documentation and lots of examples (e.g.\ the control-layer application was based on \verb|ethercrab/examples/ek1100.rs|). Interfacing with Rust is a benefit due to its ease of use, quality tooling, performance, and memory safety~\cite{RustHomePage}.

Finally, the control layer had to use the API of the middleware layer to implement the business logic. One case of that logic was to have a controller to regulate the position of each cylinder. It had to be implemented in Rust to interface with the Ethercrab library. It also required some code to initialize the EtherCAT communication and to configure the Beckhoff terminals. There's also a user interface to interactively set the cylinder position. An interesting issue came up where the position sensor signal, mapping the $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ IO signal to the range 0..1~\verb|float|, but only reached the range 0..0.5~\verb|float|. This was because the step-down resistor was larger than designed since the right size wasn't available, so the maximum sensor reading was smaller than $\qty{10}{\volt}$. This unhandled gain is an undesired interface between the Control and IO layers since it demands the controller be written with knowledge of the minutia of how the IO was implemented. Further research will go toward how to accomodate this.

Also interesting in the architecture of a control system is what development environments can be constructed for it.  The development environment is distinct from the so far discussed mechatronic stack since it is not a part of the function of the running system, but instead is the means which an engineer uses to turn an idea into assets for a mechatronic system.  For a given tech stack, any development environment generally can yield the same assets, but a good tech stack may lend itself to higher quality, more consistent output, faster development cycles, or at least to be more enjoyable for individual engineers to use.

For the control layer, I wrote the code in the Sublime Text text editor~\cite{SublimeText} and configured it to automatically format the source code according to standards for the Rust language using the \verb|rustfmt| utility~\cite{rustfmt}.  Other text editors like Emacs, Vim, or VSCodium would have also been good choices, but Sublime is familiar.  This was done on a Windows workstation, but it was compiled and run on the Raspberry Pi Linux control computer.  To achieve this, I checked the changes on the workstation into a \verb|git| version control repository and pushed them to the control computer\footnote{A small issue that comes up is that \verb|git| generally doesn't allow users to push changes to a branch of a repository that the local user has checked out, which is the case here~\cite[§receive.denyCurrentBranch]{GitConfig}.  This is to prevent data loss.  To enable this, \verb|git| can be configured to update the local branch only if there is no possibility of data loss by executing \verb|git config receive.denyCurrentBranch updateInstead|.}.  Then, though a remote terminal, I executed commands to compile and run the program.  This became quite annoying with how many steps there are, so I use the utility \verb|cargo-watch| to automatically execute the compile and run commands on the control computer when there are changes\footnote{The command was similar to \verb|cargo watch -x 'run -- eth0'|, which compiles and executes the program with the argument \verb|eth0|, which is the network interface the EtherCAT connection is on.}~\cite{CargoWatch}.  This still is somewhat tedious and so requires more development.  It also would have been mitigated by developing in some kind of Linux environment or by using a containerized environment like would be provided by Docker or Vagrant, which would have allowed running some tests on the development machine~\cite{DockerDev,VagrantDev}.  Most of the middleware was prepared with the Raspberry Pi OS image that was installed while commissioning the control computer, but there was some additional configuration, e.g.\ to install \verb|git|, that was done with the OS's \verb|apt| package management utility~\cite{apt}.  As well, library dependencies for the control program like Ethercrab were managed with the Rust \verb|cargo| utility~\cite{Cargo}.  The computer, once commissioned, did not require any development.  The IO layer's development was aided by a four-channel oscilloscope and a multimeter, seen in figure~\ref{fig:stewart_platform_demo}, to validate electrical connections and ensure signal were propagating as expected.  It was also interesting to observe the timing jitter of the program, which was generally very small, on the order of microseconds.

The in-depth consideration of the development environment was possible because of the open nature of all the tools considered.  Every tool, both a part of the control system and the development environment, have interfaces that enable and encourage users to compose or develop tools to make an environment that meets their needs.  This same discussion could not be had with regards to Dolores since its PLC-based system doesn't have interfaces for customization or external development, and any interfaces there are demand mastery of low-level communication protocols.  The development environment was simply whatever Beckhoff blessed; no more, no less.
